---
title: "advr.04 - Linear Regression"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{linreg}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Package Overview
The `advr.04` package provides an implementation of linear regression using the ordinary least squares method.
The package makes us of the Reference Class system to create a `linreg` class, which can be used for linear regression.


## Setup
```{r setup}
library(advr.04)
```

### Creating a linreg Object
To create a `linreg` object, you can use the `linreg()` function.
Given is a simple example using the built-in `iris` dataset.
```{r}
linreg_mod <- linreg$new(formula = Petal.Length ~ Species, data = iris)
```
### Methods
The `linreg` class has several methods to print information or return statistics.

#### print
The `print` method displays the formula and the calculated coefficients $\beta$.
```{r}
linreg_mod$print()
```

#### plot
The `plot` method displays two graph with "Residuals vs Fitted" and "Scale-Location".
```{r, out.width="60%", fig.height = 5, dpi=500}
linreg_mod$plot()
```

#### resid
The `resid` method returns the vector of residuals $\hat{e}$.
```{r}
head(linreg_mod$resid())
```

#### pred
The `pred` method returns the vector of predicted values $\hat{y}$.
```{r}
head(linreg_mod$pred())
```

#### coef
The `coef` method returns a named vector for the coefficients $\beta$.
```{r}
linreg_mod$coef()
```

#### summary
The `summary` method prints the summary of the linear regression with a coefficient matrix with its residual standard error $\hat{\sigma}$ and degrees of freedom $df$.
```{r}
linreg_mod$summary()
```