---
title: "ridgereg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ridgereg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Demonstration of a simple prediction workflow using the package's own `ridgereg()` function together with `caret` to compare:

- ordinary least squares,
- forward selection,
- and ridge regression (with tuning of \(\lambda\)).

Also includes an intercept-only baseline to ensure the forward-selection comparison considers models from 0 up to \(n\) covariates.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Setup

```{r setup}
set.seed(123)

library(advr.04)
library(mlbench)
library(caret)
library(leaps)
```

# Load and split data

Use the BostonHousing dataset from the `mlbench` package and split it 80/20 into training and test sets using `caret::createDataPartition()`.

```{r data-split}
data(BostonHousing)
Boston <- na.omit(BostonHousing)

trainIndex <- caret::createDataPartition(Boston$medv, p = 0.8, list = FALSE)
housingTrain <- Boston[trainIndex,]
housingTest <- Boston[-trainIndex,]

p <- ncol(housingTrain) - 1L
predictor_names <- setdiff(names(housingTrain), "medv")
```

# OLS linear regression
Linear model using all covariates.

```{r fit-ols}
fit_lm <- caret::train(
  x = subset(housingTrain, select = -medv),
  y = housingTrain$medv,
  method = "lm"
)
fit_lm
```

# Forward selection of covariates
Use caret’s `lmStepAIC` (AIC-based stepwise selection, since `leapForward` does not support intercept model) with `direction = "forward"`,
starting from the intercept-only (null) model. By specifying the scope `list(lower = ~ 1, upper = ~ .)`,
the procedure can select among models with 0, 1, ..., n predictors.
Evaluate it with 10-fold cross-validation via `trainControl(method = "cv", number = 10)`.

```{r fit-forward}
set.seed(123)
full_rhs <- stats::reformulate(termlabels = setdiff(names(housingTrain), "medv"))

ctrl_cv_fwd <- caret::trainControl(method = "cv", number = 10)

fit_forward <- caret::train(
  x = subset(housingTrain, select = -medv),
  y = housingTrain$medv,
  method = "lmStepAIC",
  trControl = ctrl_cv_fwd,
  direction = "forward",
  trace = 0,
  scope = list(lower = ~1, upper = full_rhs)
)

fit_forward
summary(fit_forward$finalModel)
```

# Ridge regression using `ridgereg()` via a custom `caret` model

Register a minimal custom caret model that calls the package's `ridgereg()` function.

Tune \(\lambda\) using 10-fold cross-validation. A log-spaced grid is often more effective for regularization paths

```{r custom-ridgereg-define}
model_ridgereg <- list(
  label = "Custom ridgereg",
  library = "advr.04",
  type = "Regression",
  parameters = data.frame(
    parameter = "lambda",
    class = "numeric",
    label = "Lambda"
  ),
  grid = function(x, y, len = NULL, search = "grid") {
    if (is.null(len) || len < 2L) len <- 15L
    if (search == "grid") {
      lam <- unique(c(0, exp(seq(log(1e-3), log(1e3), length.out = len - 1L))))
      data.frame(lambda = lam)
    } else {
      data.frame(lambda = stats::runif(len, min = 0, max = 100))
    }
  },
  fit = function(x, y, wts, param, lev, last, classProbs, ...) {
    df <- data.frame(y = y, x)
    ridge_formula <- stats::reformulate(termlabels = colnames(x), response = "y")
    ridgereg(ridge_formula, data = df, lambda = param$lambda, ...)
  },
  predict = function(modelFit, newdata, submodels = NULL) {
    newdata <- as.data.frame(newdata)
    predict(modelFit, newdata = newdata)
  },
  prob = NULL
)
```

Tune \(\lambda\) with 10-fold CV on the training set.

```{r tune-ridge}
ctrl_cv <- caret::trainControl(method = "cv", number = 10)

lambda_grid <- model_ridgereg$grid(
  x = subset(housingTrain, select = -medv),
  y = housingTrain$medv,
  len = 21, search = "grid"
)

set.seed(123)
fit_ridge <- caret::train(
  x = subset(housingTrain, select = -medv),
  y = housingTrain$medv,
  method = model_ridgereg,
  trControl = ctrl_cv,
  tuneGrid = lambda_grid
)

fit_ridge
```

Best \(\lambda\) and coefficients.

```{r best-lambda}
best_lambda <- fit_ridge$bestTune$lambda
best_lambda

coef(fit_ridge$finalModel)
```
# Test-set evaluation and comparison

Evaluate all models on the held-out test dataset.

```{r eval-test}
pred_test_lm <- predict(fit_lm, housingTest)
pred_test_forward <- predict(fit_forward, housingTest)
pred_test_ridge <- predict(fit_ridge, housingTest)

res_test_lm <- caret::postResample(pred = pred_test_lm, obs = housingTest$medv)
res_test_forward <- caret::postResample(pred = pred_test_forward, obs = housingTest$medv)
res_test_ridge <- caret::postResample(pred = pred_test_ridge, obs = housingTest$medv)

res_test_lm
res_test_forward
res_test_ridge
```

# Summary of results
On the test set, the full linear and forward-selection models perform identically

- RMSE \( \approx 4.589\)
- \(R^2\) \( \approx 0.761\)
- MAE \( \approx 3.365\)

Implying forward selection chose essentially the same predictors.

Ridge regression is marginally better on all metrics

- RMSE \( \approx 4.587\)
- \(R^2\) \( \approx 0.762\)
- MAE \( \approx 3.363\)

These gains are tiny and likely within sampling noise, but indicate that regularization can help even in this moderate-dimensional setting.